               ADDITIONAL INSTRUCTIONS FOR USING OMEN

Here are some more detailed basic instructions for using the OMEninterface, especially for opening and using programs and files. Thetopics covered are listed below.

*** Note: SOURCE CODE SAMPLE ***
*** SCREEN SCROLL ***
*** VIEWING FILES ***
*** VIEWING OMEN SYSTEM FILES ***
*** DIRECTORY SCROLL CONTROL ***
*** VIEWING DEMO DISK FILES ***
*** OPENING (RUNNING) SOFTWARE ***
*** OPENING DOCUMENT FILES ***
*** COPYING THE DEMO DISK ***
*** VISIT OMEN's HELP" MENUS ETCETERA ***


*** Note: SOURCE CODE SAMPLE ***

(This note has been placed here because it was omitted in theoriginal release notes. It is not related to the instructions thatfollow.)

Programmers should note the sample Eazy`Asm source code for theStereo`Sound`Rate`Reducer" program. It cuts the sample rate andhence the size of stereo digitized sounds in half, with some loss offidelity.

This utility program was written in under two hours. A bug foundlater was fixed and the new version saved in under five minutes. (Afew of the first demo disks went out with the bugged version, whichmade the new file size 140 bytes too short.)


*** SCREEN SCROLL ***

Atari based OMEn uses a display which is twice as tall as the actualscreen, except in TT Hirez where the screen is already two pages. Inorder to scroll down to see the bottom half of the display, simplymove the mouse to the bottom of the screen and keep going - thedisplay will scroll. To scroll back to the top, just move to the topand keep going in the same manner as going down.

The screen scroll is done with Atari's scrolling hardware. As far assoftware is concerned, the screen is actually twice as tall as whatis seen.


*** VIEWING FILES ***

To view files, a DIRectory window must be opened, by clicking onDIRs" at the top of the display. Don't move the mouse at all whenclicking, otherwise the window won't open.

Files are shown with their name followed either by a picture (icon")(for common types of files) or a four character file extension. Hereare the uses of the icons:
   Directory     Folder Icon
   Directory     Binder Icon (Binder is a special type of directory)
   Text File     Piece of Paper
   Program       Running animal (you can run" the program)
   I/O Protocol  Diamond
   I/O Port      Double Arrow
   I/O Channel mgr.  Diamond with double arrow inside
   Sound         Speaker

Initially, the directory window opens showing the Memory" rootdirectory. In the memory directory are two subdirectories: Disk`Drives" and System".


*** VIEWING OMEN SYSTEM FILES ***

To view the OMEn system files, first open a directory window (seeViewing Files, above). Click on System" without moving the mouse.The display in the window will change to show the files inside theSystem" directory, and the word System" will replace Memory" onthe second line of the window.

All OMEn's Input/ Output manager files are in System", and there arealso some application programs there. Files in System" are alreadyloaded when OMEn is started, so they will start without accessing thedisk drive.

Clicking on a program will start it running - see Opening Software",below.

To leave the System" directory, click on the word System" up nearthe top of the directory window. This will exit from the currentdirectory (System) and go out to the previous directory (Memory).


*** DIRECTORY SCROLL CONTROL ***

When there are more files in a directory than will fit in thedirectory window, the Scroll Bar along the left side of the windowbecomes active. To go up or down by a windowfull of files at a time,click on the up or down arrow keys without moving the mouse. To go upor down by just one file, click on the up or down carat symbols. Togo directly to any area of the window, click on the approximateposition desired in the vertical scroll bar underneath these symbols.Unlike GEM, the scroll bar will jump directly to that position. To dolive scrolling", hold the mouse button down in the scroll bar areaand move the mouse slowly up and down the scroll bar.
*** VIEWING DEMO DISK FILES ***

To view files on the demo disk, go back to the Memory" directory andclick on Disk`Drives. The display will change to show all the GEMdrives on your system.

Click on Drive`A (without moving the mouse). The list of the files inthe main (root") directory of the disk will be shown. At this pointthe name Disk`Drives" will be at the very top of the directorywindow, and Drive`A" will be underneath it.

Of course, a directory can be opened for any drive on the system(floppy disk B and hard drives) by clicking on that drive.

Note: The file names on disks depend on whether the file was copiedusing GEM or OMEn. OMEn uses 14 character upper and lower case filenames on GEM disks, while GEM itself uses 8 character upper case filenames. Copying an OMEn file with GEM will cause the file name to bechanged to the GEM format. To archive OMEn for BBS downloading, itwas necessary to use GEM based software, which shortened the filenames. Regrettably, the remaining 8 characters don't always describethe file very well, such as OMEn`User`Docs" being shortened toOMEN`USE".

In the main directory of the demo disk, there are several subdir-ectories shown. To find the application programs on the disk, clickon the Software" (or SOFTWARE") directory. To find the document-ation, click on OMEn`User`Docs" (or OMEN`USE") and so on. There areadditional directories for Fonts, Pictures and Sounds. The Settings"directory contains only a text file describing settings.


*** OPENING (RUNNING) SOFTWARE ***

To open (run) an application program, you must locate it (above) andthen click (without moving the mouse) on its name in the directory.The program will load and its window will open without affecting therest of the display.

By opening Text`Ed, you can view the instructions for all theindividual progams supplied. (See Opening Document Files, below).

Don't make the mistake of trying to run a program by clicking on itsinstructions in E_Util`Progs" (or E_UTILPR") documentation direc-tory instead of the program itself, which is in the Software"directory. Application programs have a running animal" icon; theOMEn text documents have a piece of paper" icon. Directories have afolder" icon.
*** OPENING DOCUMENT FILES ***

Once an application program is running, it can open documents of itsproper type. It is necessary to exit the Software" or System"directory where the program was found and go the directory where thedocuments files are to be found (See Viewing Demo DiskFiles, above).

Or, the Software" directory can be left open and the documentdirectory viewed in a second directory window, opened by clickingagain on DIRs.

When the document to be opened is located, press down the mousebutton on the file and, with the button still pressed down, move themouse into the window of the program that is to open the file. Thenlet go of the mouse button, and the file will be opened by theprogram. (With Text`Ed and text files, it is easiest to drag the fileto Text`Ed's File" button.)

Opening" a file can mean different things to different programs.Play`It`Again`Sam will play digitized sound files, while Font`Demowill make a bitmapped font for display or printout from an outlinefont definition file.

Here is a list of document file types for the demo programs. Theprograms not shown do not open document files of any type.

   PROGRAM      OPENS              FILE EXTENSION   ICON

   Text`Ed      OMEn Text files    TEXT             Piece of paper
   Paint`Booth  Bitmapped pictures BM:BMAP,BM01,... None
   Sound Player (Play`It`Again`Sam")
                Digitized Sounds   SOUN             Speaker
   Font`Demo    Outline Fonts      PS3F,OM3F,PS     None
   File`Info    Any file
   Binary`Ed    Any file (This is for programmers)
   Cheap`OMEN`TO`ASCII
   - Same as Text`Ed (changes it to TXT" for GEM)
   Stereo`Sound`Rate`Reducer
         Stereo Digitized Sounds   SOUN             Speaker


*** COPYING THE DEMO DISK ***

If the demo disk is copied from GEM, the file names are shortened toeight upper case characters and the file extensions are shortened tothree characters. Because of this, the Text`Ed icons don't work andtwo of them become duplicate file names. In addition, and moreserious, the copy of Text`Ed itself doesn't work for reasons yet tobe investigated.

The best way to copy the entire demo from GEM is to format another720K disk and do a complete disk copy by dragging the drive A icononto the drive B icon. Alternatively, a disk copy program can be usedto copy all the sectors of the disk.

From OMEn, without a hard drive, one floppy, and more than onemegabyte of memory, drag the Drive`A" directory file into theMemory" folder.
Place the copy disk in the drive. Enter the original Drive`Afolder in Disk`Drives" to access the new disk. Open the copyDrive`A" in Memory", then click on Select All" from the Dir`Misc"menu.
Drag the selected files to the directory of the new disk.
Unfortunately, the current version of OMEn is quite slow withfloppy disks.

With one megabyte or less of memory and a single floppy disk, theentire disk can't be copied at once as there isn't enough memory. (Aone meg ST with no desk accessories open holds the entire 719K demodisk, and then runs out of memory trying to open the new blankdisk--rats!) The files must be copied into Memory" in groups,transferred to the new disk, and then erased from memory to free upspace. Drive`A must be closed and re-opened each time the disk isswapped. It's easier to do the Disk Copy" from GEM.

With two floppy drives, coping the demo to a disk in Drive`B iseasier. Enter the Drive`A folder, viewing the original disk. Open asecond directory window and enter the Drive`B folder, viewing theblank disk. In Drive`A, choose Select All" from the File Misc"menu, then drag the selected files to Drive`B".

To copy the demo to a hard drive under OMEn is easiest.
Open a second directory window, and enter Drive`C". Drag theDrive`A" folder into it from the first directory window. Rename thecopy of Drive`A" (which is now on Drive`C) to OMEN" to complete theinstallation.


*** VISIT OMEN's HELP" MENUS ETCETERA ***

The instuctions in here can help get you started. There are moreinstructions in the 2.46 Release Notes which can be printed from GEM,in the Help" menus at the top of the OMEn display, and in thedocumentation files in the User`Docs directory on the demo disk. Ingeneral, the User`Docs cover the more advanced topics, while Help"and the release notes are more basic.
 